

_________________________________________________________________________________________________________________
Scientific Question:

Is increased ACE2 gene expression in lungs of current smokers correlated with an increased risk of SARS-CoV-2 infection in adult current-smokers? 

Scientific Hypothesis:

If current smokers have a higher ACE2 gene expression in lungs compared to nonsmokers, then their rate of SARS-CoV-2 infection per 100,000 compared to nonsmokers will also be higher (p-value<0.05).
Background:

ACE2 is a transmembrane protein found in epithelial and endothelial cells of the respiratory tract, lungs, kidneys, liver and heart. The SARS-COV-2 virus binds with ACE2 receptors. Smoking tobacco increases the ACE2 gene expression in the respiratory tract and lungs. 

Methods:

I have assumed a constant rate of adult smokers, meaning 13.5% of the population is smokers. And compared this with the prevalence of smoking in adults who tested positive for SARS-COV-2. Null Hypothesis: µ≤13.5
Alternate Hypothesis: µ>13.5. 
The data has been collected from a series of published studies as mentioned in the Smoking_Data.xlxs file.
__________________________________________________________________________________________________________________
Installing required packages for bio-informatics methods and data science
1)Tidyverse
2)ggplot2
3)ggpubr
3)Biobase
4)bio3d
5)BiocGenerics
6)BiocManager
7)BiocParallel
8)BioConductor
__________________________________________________________________________________________________________________
The code below installs the packages
```{r}
install.packages("tidyverse")
install.packages("ggplot2")
install.packages("ggpubr")
install.packages("Biobase")
install.packages("bio3d")
install.packages("BiocGenerics")
install.packages("BiocManager")
install.packages("readxl")
install.packages("Bioconductor")
```

__________________________________________________________________________________________________________________
Part 1: Do Smokers are more at risk for a SARS-COV-2 infection compared to non-smokers

The code below is for reading the excel data file and defining the variable Smoking_Data

```{r}
library(readxl)
Smoking_Data <- read_excel ("Smoking Data.xlsx")
```
___________________________________________________________________________________________________________________
The code below is for displaying the first 10 data points to check the data read
```{r}
head(Smoking_Data,10)# this displays the first 10 data points from the input file
```
___________________________________________________________________________________________________________________
The code below is to display the summary statistics of the data and plotting the box plot 
```{r}
summary(Smoking_Data$`% of smokers`)
library (ggpubr)
ggboxplot (Smoking_Data$`% of smokers`,ylab="% of smokers",xlab=FALSE, ggtheme=theme_minimal())# This generates a box plot for the input data
```
___________________________________________________________________________________________________________________
The code below is to draw a quantile-quantile plot of the data and the result shows that for a small sample size. Although the plot deviates slightly from normality plot, the assumption of normal distribution is viable for this data.
```{r}
library("ggpubr")
ggqqplot (Smoking_Data$`% of smokers`,ylab="% of smokers",ggtheme = theme_minimal()) # This generates a q-q plot for analysis of the distribution of the data
```
___________________________________________________________________________________________________________________
Finally, we conduct a one-sample t test for the data with a significance level of 0.05 and
Null Hypothesis: µ≤13.5
Alternate Hypothesis: µ>13.5. 
Since, the pvalue =0.02582, we reject the null hypothesis and conclude that there is sufficient evidence in favor of hypothesis that the average rate of smokers in SARS-COV-2 patients is higher than the average rate of smoking in adult population (13.5%) i.e. smokers are at higher risk of SARS-COV-2 infection that non smokers in adult human population.
```{r}
Test <- t.test(Smoking_Data$`% of smokers`, mu = 13.5,alternative="greater")# This conducts the one-sample right sided t-test. 
Test# This displays the results of the t-test
```
_____________________________________________________________________________________________________________________
Part 2: Single Cell RNA Sequencing (*Unfinished)
I had made several attempts at this but could not get it to work and could not wrap my head around how to perform this task. I have tried seeking help online and it appears there are a lot of packages available that can perform single-cell rna sequencing. 
The first problem that I faced was that the data that I have is 2.01 GB fastq file (lung sample 3e) (https://www.ncbi.nlm.nih.gov/bioproject/PRJEB4337/). I used the ShortRead package from Bioconductor to read this fastq file and then also tried reading the fastq file as a fasta file. It takes some time to load but I couldn't get any further than that. I can query the data to see the length and everything but do not understand how to use this file to perform single cell rna sequencing. Below are the packages I tried to perform the single cell rna sequencing and I am still working on this.

```{r}
if (!requireNamespace("BiocManager", quietly = TRUE))
    install.packages("BiocManager")

BiocManager::install("ShortRead")
library(ShortRead)

if (!requireNamespace("BiocManager", quietly = TRUE))
    install.packages("BiocManager")

BiocManager::install("Biostrings")
if (!requireNamespace("BiocManager", quietly = TRUE))
    install.packages("BiocManager")

BiocManager::install("msa")
library(BiocManager)
library(Biostrings)
library(seqinr)
library(msa)

library(ShortRead)
Seq_Data <- readFastq("1_130213_AH07R5ADXX_P282_111B_index6_1.fastq/P282_111B_index6_GCCAAT_L001_R1_001.fastq")

#Also tried Seq_Data <- read.fasta("1_130213_AH07R5ADXX_P282_111B_index6_1.fastq/P282_111B_index6_GCCAAT_L001_R1_001.fastq")
length (Seq_Data)

```



